Approximation as Differentiation

### 2025-04-15 appendix/notes/grothendieck.tex
mathcal for metavariables that range over categories.

\begin{definition}[Grothendieck construction for $F$]
Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
$F$ is the category where:
\begin{itemize}
\item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $C$ paired with morphisms
$\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
\end{itemize}
\end{definition}

\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
$C$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.

\begin{definition}[Category of families]
For any category $C$ define $\Fam(C)$ to be the Grothendieck construction
$\Grothendieck{\Set^{\op}}\Fam(-,C)$.
\end{definition}

\noindent In $\Fam(C)$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$ in
$\Fam(\idx{X},C)$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},C)$.

\begin{proposition}
\item
\begin{enumerate}
\item If $C$ is locally small then so is $\Fam(C)$.
\item If $C$ has binary products then so does $\Fam(C)$.
\end{enumerate}
\end{proposition}

>>>>

\begin{definition}[Grothendieck construction for $F$]
Suppose a functor $F: \cat{C} \to \Cat$. The \emph{Grothendieck construction} $\Grothendieck{\cat{C}}F$ for
$F$ is the category where:
\begin{itemize}
\item objects $X$ are pairs of an object $\idx{X}$ of $\cat{C}$ and an object $\faml{X}$ of $F(\idx{X})$;
\item morphisms $f: X \to Y$ are morphisms $\idx{f}: \idx{X} \to \idx{Y}$ in $\cat{C}$ paired with morphisms
$\faml{f}: F(f)(\faml{X}) \to \faml{Y}$ in $F(\idx{Y})$.
\end{itemize}
\end{definition}

\noindent When $F$ is contravariant, a morphism $f: X \to Y$ is a morphism $\idx{f}: \idx{X} \to \idx{Y}$ in
$\cat{C}$ paired with a morphism $\faml{f}: \faml{X} \to F(f)(\faml{Y})$ in $F(\idx{X})$.

\begin{definition}[Category of families]
For any category $\cat{C}$ define $\Fam(\cat{C})$ to be the Grothendieck construction
$\Grothendieck{\Set^{\op}}\Fam(-,\cat{C})$.
\end{definition}

\noindent In $\Fam(\cat{C})$ the objects $X$ are thus sets $\idx{X}$ paired with indexed families $\faml{X}$
in $\Fam(\idx{X},\cat{C})$ and morphisms $f: X \to Y$ are functions $\idx{f}: \idx{X} \to \idx{Y}$ paired with
morphisms $\faml{f}: \faml{X} \to \reindex{\faml{Y}}{\idx{f}}$ in $\Fam(\idx{X},\cat{C})$.

\begin{proposition}
\item
\begin{enumerate}
\item If $\cat{C}$ is locally small then so is $\Fam(\cat{C})$.
\item If $\cat{C}$ has binary products then so does $\Fam(\cat{C})$.
\end{enumerate}
\end{proposition}
